---
title: 'Exploration 1: Description in One Dimension'
author: "Jason Renn"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    number_sections: yes
    toc: yes
geometry: margin=1in
graphics: yes
mainfont: Minion Pro
fontsize: 10pt
bibliography: classbib.bib
---

<!-- Make this document using library(rmarkdown); render("exploration1.Rmd") -->
\input{mytexsymbols}



```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("exploration1.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("exploration1.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=TRUE,     # display code as typed # Changed to true because my comments are too long
  tidy.opts=list(width.cutoff=75),
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132, 
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  comment=NA)

# Packages with errors
# install.packages("psych", "robustHD")
require(psych)
require(robustHD)
```


```{r initialize,echo=FALSE}
##First, just setup the R environment for today:
if(!file.exists('figs')) dir.create('figs')

options(SweaveHooks=list(fig=function(){
			   par(mar=c(3.5, 3, 1.1, 0),
			       pty="s",
			       mgp=c(1.5,0.5,0),
			       oma=c(0,0,0,0))},
			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
			 options(prompt=" ")
			 }),
	digits=4,
	scipen=8,
	width=132
	)
options(error=function(){options(prompt="> ",continue="+ ");NULL})
```

"Brexit! UKIP! ISIL!" When your old friend calls, she seems to be yelling. Once
she calms down, she explains: "I am in charge of Improving Civic Society
programs for the United Nations, and have been asked to step in to help out at
the UK Office of Social Capital." After you congratulate her on what appears to
be a promotion she continues. "The thing is that over here in the UK, they are
really big on numbers. I asked my staff for a simple report on the status of
civic society in the UK before all of the recent unrest happened there, say, in
2005, before the London Bombings. They responded with numbers. When I asked
them to explain, I found their desks empty, their chairs knocked over, and
their computers smashed, but their coffee cups still warm and untouched." You
ask her about her own safety and she responds. "This is all within operational
parameters. No worries. My problem is that I need to report to the high command
and I don't know what the right answer is. Now I don't even have numbers. Please help. Can we hop on a
Hangout?"

She does not enable video in the Hangout. However, she begins sending you some
code. 
# "Here is what I have in terms of output. Can you explain to me what is going on?"


```{r}
# Loading a dataset in from a url
load(url("http://jakebowers.org/Data/ho05.rda"))

# Table of the 'postbomb' variable. It seems to have values of '0' and '1'. 
# Note that the option 'useNA' allows you include missing values, denoted with an "NA" value in R. There do not seem to be any missing values.
table(ho05$postbomb,useNA="ifany") 

# Here's another way to see that (explicitly) using the 'exclude' option.  
table(ho05$postbomb, exclude = NULL)

# Will use some functions from the 'dplyr' package. If you don't understand this yet, there will be a formal workshop on this material after APSA. Until then, though here is a helpful set of examples: https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html
library(dplyr) 

# You should know from the prep material that the '<-' and "=" commands are ways of assigning values in R. I suggest using the '<-' syntax, though with dplyr we will sometimes have to use an '=' when using mutate(). More on that later.
wrkdat <- ho05  %>% filter(postbomb==0)

# We see that the computer says that the sum of this variable is NA. Does that make sense? More below: 
sum(wrkdat$hlphrs)
```

_The commented lines in the code explain all but the last line, which requires that we take a closer look and execute our own code._  

```{r}
# We don't know what this variable represents, but we can look at its values by using the table command. 
table(ho05$hlphrs,useNA="ifany") 
```

_In keeping the 'useNA' option, we see that in addition to numbers ranging from 0 to 168, there are a few 'NA' values. What might you do about that? [Humm...](http://lmgtfy.com/?q=missing+values+sum()+R)_

_Okay, so google says that we can remove these values and get an actual number. Think about why this is necessary and why summing values with NAs gives an NA in return, if you're so inclined._  

```{r}
# Sum variable and remove missing values
sum(wrkdat$hlphrs, na.rm=T)
```

She continues, "And then I have this from a previous meeting where they talked about `codebooks` but I don't think these were the ordinary kind of encrypted communication behind enemy lines. [^1]: If you're compiling to html, the codebook text will look messy. HTML doesn't handle TeX well and using html would cause the opposite problem in pdf output. 

\begin{small} 
\begin{verbatim}
  ### CODEBOOK  
  postbomb: 1=interviewed after the bombing, 0=interviewed before the bombing  

  grphrs: 6.1.1 Which of the following groups, clubs or organisations  
    have you been involved with during the last 12 months? That's anything  
    you've taken part in, supported, or that you've helped in any way, either  
    on your own or with others. Please exclude giving money and anything that  
    was a requirement of your job.  

    6.1.2 In the last 12 months have you given unpaid help to any groups, clubs or  
    organisations in any of the following ways?  

    6.1.5 Approximately how many hours have you spent helping this/these group(s),  
    club(s) or organisation(s) in the past 4 weeks?  

  infhrs: In the last 12 months have you done any of the following things,  
    unpaid, for someone who was not a relative?  

    This is any unpaid help you, as an individual, may have given to other people,  
    that is apart from any help given through a group, club or organisation. This  
    could be help for a friend, neighbour or someone else but not a relative.  

    6.4.4 Now just thinking about the past 4 weeks. Approximately how many hours  
    have you spent doing this kind of thing/these kind of things in the past 4  
    weeks?  

  hlphrs: grphrs+infhrs  
\end{verbatim}
\end{small}

# She asks, "What is the best way to say how civic life was going before the bombings? What is the right answer?"

_So we have an answer for what 'hlphrs' represents. Strictly speaking, it is the sum of hours spent helping for certain organizations and individuals who are not relatives. What does it represent? Well, your friend's question suggests that it represents some form of behavior in an individual's civi life, perhaps charity. So how does this change before and after the bombings?_

```{r}
# Here is a list of 'hlphrs' values prior to the bombings. We'll use the same filter() command that we saw above, but also include the select() command so that only include one variable value
# I realize that we haven't explained the magrittr/chaining/pipe operator (%>%) yet. How about now that I've told you its many names, you investigate it? 
preBombing <- ho05 %>% 
  filter(postbomb == 0) %>% 
  select(hlphrs)
preBombing

# And post-bombing values
postBombing <- ho05 %>% 
  filter(postbomb == 1) %>% 
  select(hlphrs)
postBombing
```

_Okay, the lists are really uninformative. How about we choose a measure of central tendency to give to your friend._

```{r}
# The average, with NAs removed. What would happen if I didn't remove the NAs?
mean(preBombing$hlphrs, na.rm=T)
mean(postBombing$hlphrs, na.rm=T)
```

_What do you see? Well, it looks like the number of hours spent helping others increased after the bombing so you could tell your friend that the effect of the bombing was positive for civil life in this respect. But don't get carried away. The average number of hours increased, but we haven't talked about our confidence in this increase, whether the size is meaningful, or a number of other things._

Later, after you had worked on this a bit she calls back, "Hey. Thanks so much
for helping! I just found this code and thought it might be useful. What do you
think? 

# Can you tell me what this means? 
_Yeah, let me put some comments in the code._  

```{r, results='hide'}
# This is the average with the 10% of the values trimmed from the top and bottom. How else might you do this, say using the quantile() command or a similar approach. This would be a great Daily R activity.  
mean(wrkdat$hlphrs,trim=.1,na.rm=TRUE)

# Creates a new object out of the 'wrkdat' object where each value of hlphrs is atomized (broken down). Now each value contained in the 'hlp' vector is the number of hours and the name represents the variable name with an observation number appended to the end.  
hlp <- unlist(wrkdat %>% dplyr::select(hlphrs) %>% filter(!is.na(hlphrs)))

# In case the unlist() command still isn't making sense:  
length(hlp)
sum(!is.na(wrkdat$hlphrs))
hlp[1:5]
wrkdat$hlphrs[1:5]

# You will need to install this, probably.  
library(psych)

# Tired of google? Too bad. The simple answer to what this does: winsorizing mean. 
# This 
winsor.mean(hlp)

# What are you supposed to google? Either robustHD() or winsorize() and "R" will lead you to the documentation. Interpretation below.  
library(robustHD)
mean(winsorize(hlp))

# Here's your first user-defined function in R. 
# We will talk about this too.
onestepMest <- function(x){ # create a function named onestepMest, takes in 'x'
	## Following http://www.psychology.mcmaster.ca/bennett/boot09/rt2.pdf # You can go here to read about the method
	madn <- mad(x, constant = 1.4826) ## 1/1.4826 = .6745
	M <-  ( abs(x - median(x)) / madn )  > 1.28
	U <- sum( x > M)
	L <- sum( x < M)
	B <- length(x) - U - L
	n <- length(x)
	mux <- (1.28 * madn * (U-L) + B) /(n - L - U)
	return(mux)
}

# Interpret me. Okay, in a bit. 
onestepMest(hlp)

library(robustbase)
# Ditto, see below
huberM(hlp)
fivenum(hlp)

# Oh, hey. Turns out that the vast majority of people don't care for others at all. Well, at least not according to this measure. 50% less than 1 hour (most spent 0). At the extremes, the top ten percent spent 18+ hours and the top person spent 168, which definitely inflates the mean.  
quantile(hlp,seq(0,1,.1))

# Shoot, I want to know about this 168-hour person. I'll use indexing, if only to show you that there are different ways to do this. dplyr is cleaner. 
wrkdat[which(wrkdat$hlphrs==max(wrkdat$hlphrs, na.rm=T)),]
```

# Does it help me get the right answer about how much time people in the UK were devoting to helping each other and/or supporting groups? 
_It gives us a better idea than just the mean. It is just for the preBombing time period though, so we're not really making a comparison yet. Hey, another great Daily R assignment (or seven)_

# Why are there so many ways to descibe a single variable anyway? 
_Because everyone didn't spend the same amount of time helping people. We'd like to know if some extreme values in particular are driving the difference pre/post-bombing._

# What is the point?  
_Again, we're not even making the comparison between pre and post here, but we can be pretty sure that the average difference between the two times is misleading due to extreme values for this 'hlphrs' variable. At least the vanilla average is misleading. The others means of calculating averages, using..._

# Also, are there any plots that would help me tell the right story about this variable?
_Why yes. And if you had asked this first, we probably wouldn't have spent so much time looking at means. A picture, something-something-something words._  

```{r}
hist(preBombing$hlphrs)
hist(postBombing$hlphrs)
```


<!-- see also https://dornsife.usc.edu/labs/rwilcox/software/ and WRS:::mest  https://dornsife.usc.edu/assets/sites/239/docs/WRS2.pdf and the MASS library

This is a comment. You will not see it in the compiled document. You may use Ctrl/Cmd+Shift+C to insert a comment. Explore the other tricks and keyboard shortcuts by, you guessed it, searching for them on the internet.  

--->

# References
At some point, we will go over how to include references -- either using a bibtex or some of the citation packages. For now, feel free to paste some things in here that you used to answer the questions. They'll help show effort in addition to being good practice.  