---
title: "exploration3-questions"
author: "Duu Renn"
date: "September 9, 2016"
output:
  html_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
---


------------

/vspace{3cm}

```{r source-530-explorations, include=F}
require(knitr)
# I'm running all the R code from explorations 2 and 3 here. This way I'll have access to datasets and variables used in those assignments.
# To get this to work on your machine, you'll need to change the working directory. Look it up if you're interested!
setwd("C:/Users/duu/Documents/Dropbox/Graduate School/Fall 2016/uips-stat-share/")
source(purl("ps530/exploration2.Rmd"))
source(purl("ps530/exploration3.Rmd"))
```


```{r setup-top-level, include=F}
require(knitr)
opts_chunk$set(
  tidy=TRUE,     # display code as typed # Changed to true because my comments are too long
  tidy.opts=list(width.cutoff=75),
  size="small",    # slightly smaller font for code
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132, 
  size='footnotesize',
  out.width='.9//textwidth',
  fig.retina=FALSE,
  message=FALSE,
  warning=FALSE,
  comment=NA)
```

```{r packages, include=F, eval = F}
# install.packages(c("devtools", "knitcitations"))
require(knitcitations)
cleanbib()
options("citations_format" = "pandoc")
```

__Questions__  
As people ask questions, I'll post responses here and push them to the shared folder. You can email me these questions or create an "issue" in the 'uips-stat-share' repository. Note that this is different than the 'explorations' repository -- we should probably try to keep that relatively clutter-free.  

# Readings
Before considering non-linear relationships, make sure that you have a good understanding of linear relationships and specifically how OLS fits its line. A closer reading (and practice) with the __Kaplan__ and __James et al__ textbooks should be useful here. Chapter 3 and the "Lab" section of the __James et al__ book should be especially helpful, as it also includes some topics that are useful for this week.  

# Getting Started
Once you've read the textbook and worked through a few examplese, try returning to OLS using an example from class, but now with a continuous explanatory and outcome variable. As the exploration suggests, use the _age_ as your explanatory variable and support for Donald Trump (_fttrump_) as the outcome. You can find the codebook for this 2016 pilot of the ANES [here](http://www.electionstudies.org/studypages/anes_pilot_2016/anes_pilot_2016_CodebookUserGuide.pdf). If you don't know what a feeling thermometer is, look it up or ask an Americanist. Interested in the wording of the questions? Check [here](http://www.electionstudies.org/studypages/anes_pilot_2016/anes_pilot_2016_qnaire.pdf)/  

I put some code and comments below to help get you started


```{r}
# The variables
summary(anespi16$fttrump) # FT support for Donald Trump. Are there are surprising values?
summary(anespi16$birthyr) # Year born. Try to convert it into age.

# Fit the linear model, but exclude the 998s in the feeling thermometer
lmLinear <- lm(fttrump ~ birthyr, data=anespi16, subset=fttrump<=100)

# Guess: The greater your birthyear the lower your ft score for trump.
summary(lmLinear)
coef(lmLinear)
# Substantive interpretation
# refer to class notes

# Plot the data. Consider what line would fit? Would it fit well? What makes for a good and bad fit again, according to OLS? 
with(subset(anespi16, fttrump<=100), plot(birthyr, fttrump))
abline(a = 77, b= -.3, pch=4)

sum(lmLinear$residuals^2)

# Remeber the myssr() function? See if you can get it working for these data! A conus if you can get the contour plot working (no actual credit available).

# Here's an example of a contour plot using the ggplot package
require(ggplot2)

ggplot(data=faithfuld, aes(x=waiting, y=eruptions, z = density)) + 
  geom_contour()

```

# Null response for 'splines' 
What should the regression output look like for non-linear functions? Here's a question from a group:  

> At the moment we're working on running the smooth.spline function - although R lets us run the function, when we call up the coefficients, it results in NULL, and when we check the sum of squares of residuals, we also get 0, even when changing the degrees of freedom in our smooth.spline function.  

```{r, eval = T}
require(splines)
globalNonLinear <- smooth.spline(anespi16$fttrump, anespi16$birthyr, df=6)

# So these are what we use for linear models, but what happens when you apply them to non-linear models --- the smooth.spline() function specifically. 
coef(globalNonLinear)
sum(globalNonLinear$residuals^2)

# What is inside a lm() object?
ls(lmLinear)

# What is inside the smooth.spline object? 
ls(globalNonLinear)

```

So here are some bigger questions that might help lead to an answer. If a linear model is a slope and an intercept, what defines a spline? The [documentation](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/smooth.spline.html) may help you find the answer. How would you get residuals out of the spline model? How would you graph it?  


# Local or Global? Bias and Variance.  
As you continue to consider the exploration, think about the following:  

A) This bureaucrat discusses "local" versus "global" smoothers. What is a local smoother? What is a global one? How many ways are there to achieve smoothness? Is there anything in between? Consider the following code chunk using piecewise linear splines.  
```{r}
wrkdata <- subset(anespi16, fttrump <=100)
# linear
linearM <- lm(fttrump ~ birthyr, data=wrkdata)
# one df
fitX <- lm(fttrump ~ bs(birthyr, df = 3), data=wrkdata)
summary(fitX)


# If you have questions about for loops still, you can find an example in the shared dailyR folder
residualHolder <- as.vector(NULL)
modelObjects <- as.list(NULL)

for(i in 1:20){
  
}

sum(residuals(globalNonLinear)^2)
```

What values (there are multiple) determine the flexibility of the spline approach? To extend this, consider how you would explore the polynomial approach in the same way. Is the polynomial approach local, or global?  

B) What do James et al mean when they discuss the bias-variance trade off? How does this relate to the choice presented in this exploration? What does this mean in substantive (who votes for Trump) terms? What is the goal of the person who is asking you questions -- is it prediction or explanation?  



