---
title: "Linear Regression"
author: "Jason Renn"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 8
    highlight: haddock
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
---

\pagebreak

# Research Question
What is the outcome? What might affect this outcome? 

[Is it important?](http://elevenmyanmar.com/editorial/6521)

# Literature, Theory, Hypotheses
Jensen, Nathan M. "Democratic governance and multinational corporations: Political regimes and inflows of foreign direct investment." _International Organization_ 57.03 (2003): 587-616.

- Regime type
- Size of economy
- Conflict and stability
- Trade

```{r setup, include=FALSE}
require(knitr)
opts_chunk$set(
  tidy=TRUE,
  tidy.opts=list(width.cutoff=75),
  size="small",
  echo=TRUE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=TRUE,
  highlight=TRUE,
  width.cutoff=75, 
  fig.retina=FALSE,
  message=FALSE,
  warning=FALSE,
  comment=NA)
```

# Data
```{r, include = T, echo = T}
require(foreign)
jensen <- read.dta("~/sample-stats/workshops-tutorials/linear-regression/bookreplicationPANEL.dta")
colnames(jensen)

# Outcome: FDI Inflows (millions of USD)
summary(jensen$foreignd)

# Main Explanatory: Regime
summary(jensen$regime)

# Controls: GDP, Trade, Military Expenditures
summary(jensen$gdpperca)
summary(jensen$tradeofg)
summary(jensen$military)
```

# Bivariate Association
```{r, include = T, echo = T}
require(ggplot2)
ggplot(data=jensen, aes(x=regime, y=foreignd)) +
  geom_point()

ggplot(data=jensen, aes(x=regime, y=foreignd)) +
  geom_point() + 
  stat_smooth(method = "lm")

ggplot(data=jensen, aes(x=regime, y=foreignd)) +
  geom_point() + 
  stat_smooth(method = "lm") + 
  scale_y_continuous(limits = c(-50, 100))
```

# Linear Regression
```{r, results='asis'}
lm1 <- lm(foreignd~regime, data=jensen)
require(stargazer)
stargazer(lm1, header = F)
```

## Coefficient
Meaning and substantive interpretation 

## Standard Error and P-values
```{r}
summary(lm1)
```

## Model Fit and Control Variables
```{r, results='asis'}
lm2 <- lm(foreignd~regime + gdpperca + tradeofg + military, data=jensen)
stargazer(lm2, header = F)
```

```{r}
summary(lm2)
```

```{r, results='asis'}
lm3 <- lm(foreignd~regime + gdpperca + tradeofg + military +rnorm(nrow(jensen)), data=jensen)
```

```{r}
summary(lm3)
```

## Association of Cause? 



